---
title: "Sholl analysis consolidation"
author: "James Crichton"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sholl analysis 

This is an R-markdown script to consolidate, visualise and analyse data generate by sholl analysis (ImageJ) of imaged segmented from brain tissue sections using QuPath.

## Step 1: read in data

User must define the input directory containing the image data. This should be generated from previous scripts, and posess the title "exported for sholl/". Subdirectories will be 1) Image name; 2) cell name.

Everything else beyond this cell should be automated and require no direct user input.

```{r data input, include=FALSE}

# User should input the QuPath project's "exported_for_sholl" directory, generated in earlier steps of the analysis
dir <- "Q:/Kate Ellacott Lab/Alex Mellor/JC Analysis Support/Exploratory QP proj/exported_for_sholl" 

```

```{r dependencies, echo=FALSE}

# Load libraries
library(tidyverse)


# Consolidate data into a single df
df_joined <- data.frame()

all_files <- list.files(path = dir, recursive = TRUE, full.names = TRUE)
csv_files <- all_files[grepl("\\.csv$", all_files)]

counter = 1

for(path in csv_files){
  path_levels <- strsplit(path, .Platform$file.sep)[[1]]
  
  cell_df <- read.csv(path)
  cell_df$cell <- path_levels[length(path_levels)-1]
  cell_df$img <- path_levels[length(path_levels)-2]
  cell_df <- cell_df %>% select(!X.)  # Drop this empty col

  df_joined <- bind_rows(df_joined, cell_df)

  # print(paste(counter, "of", length(csv_files), "complete", sep=" "))
  # counter = counter + 1
}
```

## Plot data

```{r}

# Plot data from all cells
rad_int_all_lineplot <- ggplot(df_joined)+
  geom_line(aes(x=Radius, y=Inters., col=cell))+
  ggtitle("All cells") +
  theme(plot.title = element_text(hjust = 0.5))

rad_int_all_lineplot


# Plot log.Inters..Area. data agains radius, after filtering for empty rings
rad_logintarea_all_lineplot <- 
  df_joined %>% 
    filter(log.Inters..Area.!=0) %>%  # Filter zero values (no intersections) as these are not meaningful)
    ggplot()+
      geom_line(aes(x=Radius, y=log.Inters..Area., col=cell))+
      ggtitle("All cells") +
      theme(plot.title = element_text(hjust = 0.5))

rad_logintarea_all_lineplot


# Pooled data with sd shown in the opaque area behind line
pooled_sholl_plot <-  df_joined %>% 
    filter(log.Inters..Area.!=0) %>%  # Filter zero values (no intersections) as these are not meaningful) %>%
    group_by(Radius, img) %>%
    summarise(intersections_mean = mean(Inters.), 
              intersections_sd = sd(Inters.)) %>%
    mutate(sd_upper = intersections_mean+intersections_sd, 
           sd_lower = intersections_mean-intersections_sd) %>%
    ungroup() %>%
    ggplot(aes(x=Radius, y=intersections_mean))+
      geom_line()+
      geom_ribbon(aes(ymin=sd_lower, 
                      ymax=sd_upper),
                  fill = "blue", alpha = 0.2)+
      ggtitle("Pooled sholl analysis results\nBlue ribbon = st dev") +
      theme(plot.title = element_text(hjust = 0.5))+
  xlab("Radius from centroid (px)")+
  ylab("Mean intersections")
    
pooled_sholl_plot
    
```
